---
title: "Covid Analysis Report"
date: "2025-04-25"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(ggplot2)
library(dplyr)
library(knitr)
```

## Introduction

This report summarizes findings from an analysis of COVID-19 patient data in Mexico. The dataset includes demographic and clinical features of over a million patients, with a 20% subset used for initial analysis. We focused on how age and pre-existing health conditions relate to COVID-19 mortality using descriptive visualizations and logistic regression modeling. All results in this report were generated ahead of time and saved in the `output/` folder for reproducibility.

## COVID-19 Mortality by Age Group

```{r, echo=FALSE, out.width="80%"}
knitr::include_graphics(here("output", "figures", 
                             "bar_age_group_vs_death.png"))
```
This bar chart illustrates the relationship between patient age group and COVID-19 mortality outcome. Younger age groups (<18, 18–39) had very low death proportions, whereas mortality increased significantly in the 40–59 group and was highest among individuals aged 60 and above. These results highlight age as a critical factor in predicting COVID-19 severity.

## Logistic Regression Model of Mortality Risk

```{r model_results, echo=FALSE}
glm_coef <- readRDS(here::here("output", "tables", 
                               "glm_coefficients.rds"))

# Round numeric columns to 2 decimal places
glm_coef_clean <- glm_coef %>%
  select(-conf.low, -conf.high) %>%
  mutate(across(where(is.numeric) & !matches("p.value"), ~ round(.x, 2))) %>%
  mutate(p.value = ifelse(p.value < 0.001, "<0.001", round(p.value, 3)))

kable(glm_coef_clean, caption = "Logistic Regression Coefficients for COVID-19 Mortality")
```

We fitted a logistic regression model to predict the likelihood of death (death_indicator) based on demographic and clinical factors. Key findings:

Age has a strong positive association with mortality (p < 0.001).

Chronic conditions such as diabetes, hypertension, immunosuppression, obesity, and renal disease significantly increase the risk of death.

Asthma was negatively associated with death (p < 0.001).

Tobacco use and cardiovascular disease were not significant predictors in this model (p > 0.05).